Chronic conditions and increased antibiotic consumption

### 2017-03-14 chronic-conditions/chronic-condition-drivers.Rmd
chronic: lasso

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv')
chronic_conditions = read_tsv('cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG)
condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
```

```{r summarize_users}
usage_summary = usage %>%
  filter(age >= 65) %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims), age=max(age), sex=head(sex, 1)) %>%
  mutate(is_female=if_else(sex=='female', 1L, 0L)) %>%
  select(-sex) %>%
  left_join(chronic_conditions, by='bene_id')
```

```{r compute_correlations}
tetra = chronic_conditions %>%
  select(-bene_id) %>%
  mutate_all(function(x) x == 3) %>%
  tetrachoric()
```

```{r compute_models}
conditions = c("AMI", "ALZH", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")

library(broom)
summarize_condition = function(x, condition) {
  has_condition = x[[condition]] == 3

  x %>%
    glm(n_claims ~ has_condition, family='poisson', data=.) %>%
    tidy %>%
    mutate(condition=condition)
}

@@

```{r multivariate_model}
multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>%
  formula

multivariate_res = usage_summary %>%
  lm(multivariate_formula, family='poisson', data=.) %>%
  tidy
```

@@

### Multivariate

I also performed a multivariate Poisson regression, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.

@@

```{r cache=FALSE}
multivariate_res %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  arrange(desc(pct.increase)) %>%
  kable
```

>>>>

```{r load_data}
usage = read_tsv('../bene_usage_2011.tsv')
chronic_conditions = read_tsv('cc_2011.tsv') %>%
  rename(bene_id=BENE_ID) %>%
  select(-STRCTFLG) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) x==3)
condition_names = read_tsv('names.txt', col_names=c('condition', 'condition_name'))
```

```{r summarize_users}
denom = usage %>% select(bene_id, age, sex) %>% distinct

usage_summary = usage %>%
  filter(age >= 65) %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  left_join(denom, by='bene_id') %>%
  mutate(is_female=sex=='female') %>% select(-sex) %>%
  left_join(chronic_conditions, by='bene_id')
```

```{r}
chronic_conditions %>% head(100) %>% write_tsv('tmp')
```

```{r compute_correlations}
tetra = chronic_conditions %>%
  select(-bene_id) %>%
  tetrachoric()
```

```{r univariate_models}
conditions = c("AMI", "ALZH", "ALZHDMTA", "ATRIALFB", "CATARACT", "CHRNKIDN", "COPD", "CHF", "DIABETES", "GLAUCOMA", "HIPFRAC", "ISCHMCHT", "DEPRESSN", "OSTEOPRS", "RA_OA", "STRKETIA", "CNCRBRST", "CNCRCLRC", "CNCRPRST", "CNCRLUNG", "CNCRENDM", "ANEMIA", "ASTHMA", "HYPERL", "HYPERP", "HYPERT", "HYPOTH")

library(broom)
summarize_condition = function(x, condition) {
  univariate_formula = paste0('n_claims ~ age + is_female + ', condition) %>%
    formula

  x %>%
    glm(univariate_formula, family='poisson', data=.) %>%
    tidy %>%
    mutate(condition=condition)
}

@@

```{r lasso}
library(glmnet)
lasso_x = usage_summary %>%
  select(age, is_female, AMI:HYPOTH) %>%
  mutate_at(vars(AMI:HYPOTH), function(x) if_else(x==3, 1L, 0L)) %>%
  as.matrix
lasso_y = usage_summary$n_claims
lasso_model = glmnet(lasso_x, lasso_y, family='poisson')
```

```{r multivariate_linear_model}
multivariate_formula = paste(c('n_claims ~ age + is_female', conditions), collapse=' + ') %>%
  formula

multivariate_linear_res = usage_summary %>%
  lm(multivariate_formula, data=.) %>%
  tidy
```

```{r multivariate_poisson_model}
multivariate_poisson_res = usage_summary %>%
  glm(multivariate_formula, family='poisson', data=.) %>%
  tidy
```

## Summary

`r usage_summary %>% kable`

@@

### Multivariate models

I also performed multivariate regressions, keeping all the chronic conditions as well as age and sex (coded as "is female") as explanatory variables.

#### Multivariate linear

I don't expect a linear model to work well, since the output is count data.

@@

```{r cache=FALSE}
multivariate_linear_res %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  arrange(desc(pct.increase)) %>%
  kable
```

#### Multivariate Poisson

This should be more accurate.

```{r cache=FALSE}
multivariate_poisson_res %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100)) %>%
  arrange(desc(pct.increase)) %>%
  kable
```

#### Lasso

I also performed a lasso regression with the Poisson model. I'm not sure how to display the results sensibly. I picked a value of $s$ that limited me to just a few nonzero model coefficients.

```{r cache=FALSE}
lasso_model %>%
  coef(s=0.25) %>%
  tidy %>%
  select(term=row, coefficient=value) %>%
  kable
```
